---
title: 'Lab #6'
output: html_document
date: "2024-02-28"
---

## Introduction:

To begin the lab 6 tasks, I first loaded the pacman library and went ahead to load othe required libraries. Next, I imported my 3 dataset which I will be making use of for this lab (Food Atlas, US State Boundary and US 2010 Census Tract dataset). I tidied up my dataset by removing variables which I will not be using in the analysis from the food atlas dataset. Looking at the structure of my data, I have 2 spatial data (US states and US 2010 Census tract) and one non-spatial data (Food Atlas). Seeing as I will be mapping the food accessibility across census tracts, I joined the census tract spatial data with the non spatial data,using the GEOID10 variable, to create a new spatial data object suitable for the task. I also created a new character variable using the binary values in LAhalfand10 variable to stand for Low acess and not low access. Finally, I created a new spatial data object with the four states that will be used for analysis. 

*I initially downloaded the wrong census data and when through a tough time trying to create avariable with which I could join the spatial US state data and the other data objects and because it took me a lot of time to figure it out, I thought to leave the code here so you could see it.*

```{r warning = FALSE, message = FALSE, error = FALSE}
library(pacman)
p_load(sf, tidyverse, ggspatial, tmap, spData)

FoodAtlas <- read_csv("FoodAtlas.csv")

foodAtlas_main <- FoodAtlas %>% 
  select(-(25:134))


data("us_states")
US_State1 <- us_states %>% filter(NAME %in% c("Kansas", "Missouri", "Oklahoma", "Nebraska"))


US_2010_CT <- st_read("US_tract_2010.shp")


# str(US_2010_CT)
# str(foodAtlas_main)

foodAtlas_sf <- left_join(US_2010_CT, foodAtlas_main, 
                          by = c("GEOID10" = "CensusTract")) %>% 
  mutate(FoodAccess = ifelse(LAhalfand10 == 1,"Low Access", "Not Low Access"))

StateBoundary <- foodAtlas_sf %>% filter(State == c("Kansas", "Missoiri", "Oklahoma", "Nebraska"))

st_crs(StateBoundary)




# us_2010_CT_new <- US_2010_CT %>% 
#   select_if(~any(!is.na(.))) %>% 
#   mutate(CT = str_remove(GISJOIN, "G"),
#          CT = paste0(substr(CT, 1, 2), substr(CT, 4, nchar(CT)))) %>%
#   mutate(CensusTract = paste0(substr(CT, 1, 5), substr(CT, 7, nchar(CT))))

```



### Question 1:

Using ggplot and sf, I created a map showing food accessibility across Census Tracts for the states of Kansas, Missouri, Oklahoma, and Nebraska. There were lots of NA values in the dataset and no values in Missouri. 

```{r warning = FALSE, message = FALSE, error = FALSE}
StateBoundary %>%
  ggplot() +
  geom_sf(aes(fill = FoodAccess), color = "white", size = 0.2) +
  scale_fill_viridis_d(name = "Food Accessibility") +
  labs(title = "Food Accessibility Across Census Tracts",
       subtitle = "States: Kansas, Missouri, Oklahoma, Nebraska") +
  theme_minimal() +
  theme(axis.text = element_blank(), axis.title = element_blank(), legend.position = "right") +
  geom_sf(data = US_State1, fill = NA, color = "black", size = 1) +
  geom_sf_text(data = US_State1, aes(label = NAME, position = "jitter", label.size = 0.01)) +
  theme(axis.text = element_blank())

```


### Question 2:

For the state of Kansas, I generated a map to show the distribution of the low income tract (TractLOWI) based on food accessibility.

```{r warning = FALSE, message = FALSE, error = FALSE}

TLI_pt <- st_centroid(StateBoundary)

StateBoundary %>%
  filter(State == "Kansas") %>%
  ggplot() +
  geom_sf(aes(fill = FoodAccess), color = "grey", alpha = 0.7) +
  geom_sf(data = TLI_pt %>% filter(State == "Kansas"), 
          aes(size = TractLOWI)) +
  scale_fill_viridis_d(name = "Food Accessibility") +
  scale_size_continuous(name = "Low Income Tract", guide = "legend") +
  labs(title = "Food Accessibility and Low Income Tracts Across Census Tracts in Kansas") +
  theme_minimal() +
  theme(axis.text = element_blank(), axis.title = element_blank(), 
        legend.position = "bottom",
        legend.key.size = unit(0.5, "cm")) +
   ggspatial::annotation_scale(
    location = "bl",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book") +
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0.4, "in"), pad_y = unit(0.4, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"))
```


### Question 3:

Here, I created 3 seperat maps using base R plot(), ggplot and tmap to show the prevalence of low food accessibility across the four US stateswhich I have been working on i.e. the state of Kansas, Oklahoma, Missouri, and Nebraska.

```{r warning = FALSE, message = FALSE, error = FALSE}
#Using base R to plot

plot(StateBoundary[ ,"FoodAccess"], nbreaks = 3, breaks = "pretty",
     col = c("Not Low Access" = "darkgreen", "Low Access" = "darkred"),
     main = "Food Accessibility across 4 US States",
     key.pos = 1)
legend("bottomright", legend = c("Not Low Access", "Low Access"), fill = c("green", "darkred"))



#Using ggplot

StateBoundary %>% 
  ggplot() +
  geom_sf(aes(fill = FoodAccess), alpha = 0.7, col="black") +
  labs(title = "Prevalence of low food accessibility across US states") +
  theme_minimal() +
  theme(axis.text = element_blank(), legend.position = "right") +
  guides(fill = guide_legend(title = "Food accessibility \nacross US states", 
                             title.size = 0.5)) +
  scale_fill_viridis_d(option="magma", begin=0.25, na.value = "grey")+
   geom_sf(data = US_State1, 
          fill = NA, color = "black", size = 1) +
  geom_sf_label(data = US_State1, aes(label = NAME), label.size = 0.005, label.padding = unit(0.005, "lines"))



#Using tmap

tmap_mode("plot")

StateBoundary %>%
  qtm(fill = "FoodAccess",
       fill.title = "Food accessibility\nacross US states",
       legend.outside.position = "bottom",
       title = "Prevalence of low food accessibility across US states", 
       style = "col_blind") +
  tm_layout(main.title.size = 3.5, 
            main.title.color = "darkblue",
            frame = FALSE,
            inner.margins = c(0.1, 0.1, 0.1, 0.1)) +
  tm_borders(lwd = 0.5, col = "grey") +
  tm_legend(outside.position = "bottom", 
            text.size = 0.8, 
            bg.color = "white")+
  tm_compass() +
  tm_scale_bar()


```



### Question 4: 

Finally, I created an interactive map showing information on food accessibility and Median Family income across counties using Stadia.AlidadeSmoothDark as the basemap. When the cursor hovers above each bubble, it give a number 40 and althoght I tried to figure out why, I couldn't.


```{r warning = FALSE, message = FALSE, error = FALSE}
#Make interactive map

tmap_mode("view")


tm_basemap("Stadia.AlidadeSmoothDark") + 
  tm_shape(StateBoundary  %>% filter(State == "Oklahoma")) +
  tm_borders(lwd = 0.5, col = "white") +
  tm_bubbles(size = "MedianFamilyIncome", col = "steelblue", 
             popup.vars = c("LAhalfand10", "FoodAccess")) +
  tm_text("County", size = 1, root = 4, col = "black")
```

